---
title: "gasmet_data-conversion_2022"
author: "jesse puka-beals"
date: "2022-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Overview

Gasmet measurements are recorded using calcmet software

This software generates .SPE files that are autosaves into a folder named by date. For example folder 20221027 is for the most recent sampling on October 27th. Each time someone takes measurements on a new day, the data from the last sampling date is compressed. For example, there was once a folder called 20221021 but it was compressed into "SAMPLE20221021_1" on october 27th when new samples were recorded. 

Our objective is to get all of the data into a single data frame that has the date, the experiment, the field, the plot and the predicted gas concentrations (i.e. CO2 in ppm). 

# Scope

only 2022 data

# Workflow steps

1. convert the cabinet files (a compressed file format) into an unzipped and uncompressed format

2. load the SPE files into calcmet and have calcmet predict the gas concentrations of interest and output those predictions into large files

3. combine those large files of predicted values into a single file

4. tidy that large file 

5. done

# convert cabinet files

Cabinet files are located in the SEECRS google drive under gasmet data and then 2022

I will make a new folder on my PC and copy all those files onto my physical computer first, in case there are any issues during this process, we are working on copies. 

I found a free software called 7-Zip that converts/unzips cabinet files. At the time, my windows operating system could not do it

Now I simply navigate within Z-zip to the folder on my PC that contains the cabinet files and select "extract". This takes about 2 minutes. To make this easier, I select "no pathnames" so it doesn't preserve the folder structure and I create a new folder called "data_all_spe" where I tell 7-Zip to dump all the files

Now I drag in the files from the one sampling day that wasn't compressed 

I also copy in the SPE files from that one folder that was not compressed "20221027".

Now all the SPE files are in one place for 2022

# load SPE files into calcmet and generate outputs

Within Calcmet, I go to the toolbar on the top and select "Options" then "Result Output". 

I make a new folder and send the result outputs to this new folder. They will all be in one file so I name the file "batch1"

There are 20,629 files we are sending through gasmet. 

# combine large output files into single file

I tried in 2021 and again in 2022 to drag all SPE files into Calcmet and then generate one large output file, but it seems to be unable to produce a file larger than ~5k observations, so I have to break up the files and run them through calcmet in batches

```{r}

all_spe <- "C:\\Users\\pukab001\\Documents\\gasmet_data\\spe_all_2022\\"

list.files(all_spe) %>% 
  length()

20629/7


```

We can create 7 folders and send 2947 files to each folder without any leftover

```{r}
basepath <- "C:\\Users\\pukab001\\Documents\\gasmet_data\\"
paste0(basepath,"batches")
dir.create(paste0(basepath,"batches"))

for(i in 1:7){
  dir.create(paste0(basepath,"batches\\","batch",i))
}

```

we will need a key to tell the loop where to start and end it's indexing

```{r}
tibble(
  batch = 1:7,
  start = seq(from=1,by=2947,length.out=7),
  end = start+2946
) -> loop_key
```


```{r}

list.files(all_spe)[1]
paste0(all_spe,list.files(all_spe)[i])

for (i in 1:length(loop_key$batch)) {
  file.copy(
  from = paste0(all_spe,list.files(all_spe)[loop_key$start[i]:
                                            loop_key$end[i]]),
  to = paste0(basepath,"batches\\","batch",i)
)
}

```

Now we drag these files into calcmet.

First I go into Calcmet and select Options and Result Output. I change the file name each time to reflect the batch. For example, I enter batch1 for the result output file name, then I drag all of the SPE files into Calcmet and the generate the txt file output. Then I change the result output filename to batch2, then drag all the files from batch2 into calcmet and so on. 


**combine output files into a single file**

```{r}

path <- "C:\\Users\\pukab001\\Documents\\gasmet_data\\calcmet_outputs\\"

list.files(path)

paste0(path,list.files(path)[1])

# read.delim(paste0(path,list.files(path)[1]))
# multibyte string error, weird
# after troubleshooting for 1hr, solution is below fileEncoding arg
# unclear if this is due to update with R or Calcmet
# https://stat.ethz.ch/pipermail/r-help//2014-April/373464.html

read.delim(paste0(path,list.files(path)[1]),
                  fileEncoding="ISO-8859-1") -> all_batches

for (i in 2:7) {
        next_batch <- read.delim(paste0(path,list.files(path)[i]),
                  fileEncoding="ISO-8859-1")
        all_batches <- rbind(all_batches,next_batch)
}
```

# tidy large file

Now we have one large file!

Time to clean up some column names

```{r}
dat <- all_batches
View(dat)

dat1 <- dat %>%
  dplyr::select(Date,Time,SpectrumFile,Carbon.dioxide.CO2,
                Nitrous.oxide.N2O,Methane.CH4,
                Ammonia.NH3,Cell.temperature) %>%
  rename(date=Date,
         time=Time,
         co2=Carbon.dioxide.CO2,
         spectrumfile=SpectrumFile,
         n2o=Nitrous.oxide.N2O,
         ch4=Methane.CH4,
         nh3=Ammonia.NH3,
         temp=Cell.temperature) %>% 
  rowid_to_column()

View(dat1)

```

Now we want to split out that Spectrum file name

The root for all these batches is the same so this will be super easy in the instances where the naming convention was followed

```{r}
library(stringr)

# so all these files are from the same folder so this will
dat1$spectrumfile[1]

str_sub(dat1$spectrumfile,56,-5)[1]

dat1 %>% 
  mutate(filename = str_sub(spectrumfile,56,-5),
         .before=date) -> dat2
```


```{r}


dat2$filename[1]
str_split(dat2$filename,
          pattern = "_",
          simplify = T) %>% 
  as.data.frame() %>% 
  dplyr::rename(experiment = V1,
                site = V2,
                date_check = V3,
                plot = V4, 
                calcmet_chronological_tally = V5) %>% 
  cbind(dat2) -> dat3

```

Find the trouble makers

So some of these start with "R_" and we need to get rid of those!

```{r}

dat3 %>% 
  distinct(experiment)

# starting with "R"
dat3 %>% 
  filter(experiment == "R") %>% 
  mutate(filename = str_sub(filename,3)) %>% 
  dplyr::select(-c(1:7)) -> dat_Rfix
# correct file names

str_split(dat_Rfix$filename,
          pattern = "_",
          simplify = T) %>% 
  as.data.frame() %>% 
  dplyr::rename(experiment = V1,
                site = V2,
                date_check = V3,
                plot = V4, 
                calcmet_chronological_tally = V5) %>% 
  bind_cols(dat_Rfix) -> dat_Rfix2

# pull row id's of the new good dataset
dat_Rfix2 %>% 
  pull(rowid) -> dummy

# filter out all the rowid's from the dat3 dataset that we will replace, then replace

dat3 %>% 
  # colnames()
  # .["rowid" == 5]
  filter(!rowid %in% dummy) %>% 
  bind_rows(dat_Rfix2) -> dat4

```

```{r}
dat4 %>% 
  filter(experiment!="ZERO") -> dat5
```

When people used underscores to seperate month and day and year, it ruined everything!!!

```{r}
dat5 %>% 
  # distinct(experiment)
  mutate(V7 = na_if(V7,"")) %>% 
  drop_na(V7) %>% 
  dplyr::select(-c(1:7)) -> dat6

dat6$filename[1] %>% 
  str_sub(10,19) %>% 
  str_replace("_","-") %>% 
  str_replace("_","-")

dat6 %>% 
  # slice(1:3) %>% 
  mutate(date1 = str_sub(filename,10,19),
         date1 = str_replace(date1,"_","-"),
         date1 = str_replace(date1,"_","-")) %>% 
  # dplyr::select(date1)
  mutate(start = str_sub(filename,1,9),
         end = str_sub(filename,20)) %>% 
  mutate(filename = paste0(start,date1,end)) %>% 
  dplyr::select(-c(date1,start,end)) -> dat_datefix

str_split(dat_datefix$filename,
          pattern = "_",
          simplify = T) %>% 
  as.data.frame() %>% 
  dplyr::rename(experiment = V1,
                site = V2,
                date_check = V3,
                plot = V4, 
                calcmet_chronological_tally = V5) %>% 
  bind_cols(dat_datefix) -> dat_datefix2

# working back into dat5
dat5 %>% 
  filter(!rowid %in% dat_datefix2$rowid) %>% 
  bind_rows(dat_datefix2) -> dat7
  
```

What is going on with this weird eddyflux name

```{r}
dat7 %>% 
  # distinct(experiment)
  # filter(experiment=="eddyfluedddyflux")
  mutate(experiment = fct_recode(experiment,
                                 "eddyflux" = "eddyfluedddyflux")) %>% 
  dplyr::select(-c(V6,V7))-> dat8

```

```{r}
dat8 %>% 
  # colnames()
  filter(experiment!="till") %>%  # this is 2021 data
  filter(plot!="test" & plot!= "trial") -> dat9
  
# some plot numbers have an added "b" like 202b. I will leave these in the datasets
```

```{r}
dat9 %>% 
  # filter(co2>300) %>% 
  ggplot(aes(co2)) +
  stat_bin(aes(fill=experiment))

dat9 %>% 
  group_by(date) %>% 
  summarise(median = median(co2)) %>% 
  arrange(median)
# looks like a lot of september/october data is bad

dat9 %>% 
  filter(co2>300) %>%
  tally()

(20391-14617)/20391*100
# lost 28% of our dataset to the gasmet not recording measurements

dat9 %>% 
  filter(co2>300) -> dat10

dat10[duplicated(dat10$rowid),]
# no duplicated rowid values

```


```{r}
dat10 %>% 
  mutate(experiment = str_to_lower(experiment)) %>% 
  rename(co2_ppm = co2,
         n2o_ppm = n2o,
         # n2o_ppm = n20,
         ch4_ppm = ch4,
         nh3_ppm = nh3,
         temp_celsius = temp
         ) -> dat11
  
dat11 %>% 
  write.csv(
    file = "gasmet_2022_all-site-data.csv",
    row.names = F
  )
```

# parsing it out 

```{r eval=F}

dat11 %>% 
  # distinct(experiment)
  # filter(experiment=="atat") %>% 
  # filter(experiment=="eddyflux") %>%
  # filter(experiment=="fert") %>%
  filter(experiment=="ncas") %>%
  write.csv(
    file = "gasmet_2022_ncas.csv",
    row.names = F
  )

```

# Data vis

Time to enjoy the fruits of our labor

```{r}
dat11 %>% 
  ggplot(aes(co2_ppm)) +
  stat_bin(aes(y=after_stat(density))) +
  geom_density(aes(fill=date),
               alpha=.3,
               col=NA) +
  theme_bw()

# ggsave("co2_dist.png",
#        width = 6,
#        height = 6,
#        units = "in",
#        dpi=400)

```

```{r}
dat11 %>% 
  ggplot(aes(n2o_ppm)) +
  # stat_bin(aes(y=after_stat(density))) +
  geom_density(aes(fill=date),
               alpha=.3,
               col=NA) +
  theme_bw()

dat11 %>% 
  ggplot(aes(ch4_ppm)) +
  # stat_bin(aes(y=after_stat(density))) +
  geom_density(aes(fill=date),
               alpha=.3,
               col=NA) +
  theme_bw()

dat11 %>% 
  ggplot(aes(nh3_ppm)) +
  # stat_bin(aes(y=after_stat(density))) +
  geom_density(aes(fill=date),
               alpha=.3,
               col=NA) +
  theme_bw()

dat11 %>% 
  ggplot(aes(temp_celsius)) +
  # stat_bin(aes(y=after_stat(density))) +
  geom_density(aes(fill=date),
               alpha=.3,
               col=NA) +
  theme_bw()


```

```{r}
dat11 %>% 
  mutate(date = as.POSIXct(date)) %>% 
  ggplot(aes(date,co2_ppm)) +
  geom_jitter(width=.1,
              alpha=.1) +
  stat_smooth(method = "lm",
              se=F,
              linetype=3,
              col=1) +
  labs(x="",
       caption = "overall respiration seemed to decline later in year, drought?")
```

